Parsing Postfix log files

### 2008-08-25 doc/logparser-results.tex
Dealt with some XXX items: * Mentioned prior art. * Added a little more about detecting overlapping regexes. * Removed "Why are there dips in the graphs?" paragraphs, they weren't   particularly important any more. * Changed normal to optimal.
Changes made following suggestions from diction.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day; \graphref{Mails received per day} shows that as expected there are far
more mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \SMTP{} only; in particular the mail loops noticeable in
other graphs do not contribute to these figures.

@@

\Graphref{parsing time vs file size vs number of lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline
\tab{}\texttt{$\backslash$username, username@domain} \newline This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this case is the address the mail was originally sent to,
creating an infinite loop.  To prevent this happening Postfix examines the
Delivered-To header in the mail, and if the mail has already been delivered
to the current address it is bounced back to the sender with the error
message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Ordinarily this
works well, but unfortunately in this case the user noticed they had not
received any mail in a while and opted to send a test mail to themselves,
causing a loop not caught by Postfix:

@@

    \item Postfix accepts the mail for username@domain, but while
        delivering it notices that the \texttt{Delivered-To} header already
        contains the address it is currently delivering to, and therefore
        sends a bounce notification with sender address \texttt{<>} to
        the original sender: username@domain.

@@

    \item Postfix accepts the forwarded bounce notification but while
        delivering the mail it notices that the \texttt{Delivered-To}
        header already contains the address currently being delivered to,
        and sends a bounce notification to the sender: username@domain.

@@

The different distribution of log lines during a mail forwarding loop has
interesting effects on parsing time, described in detail in
\sectionref{rule ordering for efficiency}, \sectionref{scalability as the
number of rules rises}, and \sectionref{Caching compiled regexes}.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% match the vast majority of log lines, with the remaining log lines
split across the other 90\% of the rules (as shown in \graphref{rule hits
graph}).  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently matching rules before those which match less frequently.  To
test this hypothesis three full test runs were performed with different
rule orderings:

@@

    \item [optimal]  The most optimal order, according to the hypothesis:
        rules which match most often will be tried first.

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components which have a small number of rules associated with them, whereas
in general the majority of log lines are from Postfix components which have
a high number of associated rules; this results in the average number of
rules required to parse a log line from log files 22 or 62--68 being
considerably lower than usual.  The ratio of number of lines versus parsing
time in \graphref{parsing time vs file size vs number lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows much less difference
between the orderings at log files 22 and 62--68.

@@

    \item Check if there are any rules for the Postfix component which
        produced the log line; if not then skip the log line.

@@

    \item Each Postfix component whose log lines are of interest must have
        at least one rule or its log lines will be silently skipped; in the
        extreme case of zero rules the parser would happily skip every log
        line.  There may be any number of log lines from other programs
        intermingled in the log file, and there are some Postfix programs
        which do not produce any log lines of interest.

    \item There must be a rule to match each different log line produced by
        each program; if a log line is not successfully matched the parser
        will issue a warning.  Rules should be as specific and tightly
        bound as possible to ensure accurate parsing:\footnote{A rule which
        matches zero or more of any character will successfully parse every
        log line, but not in a meaningful way.} most log lines contain
        fixed strings and have a rigid pattern, so this is not a problem.

@@

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6,039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.\footnote{Each log
line was examined and the correct \regex{} identified from the
\numberOFrules{} rules in the database; the correct \regex{} was then
compared to the \regex{} which was used by the parser.}  The sample was
generated by running the following command:

@@

\noindent{}to randomly extract roughly one line in every 10,000.  Although
on initial appearances exercising only 36 rules (from a total of
\numberOFrules{}) when parsing 6039 log lines seems quite low, after
examining \graphref{rule hits graph} it becomes apparent that such
a low hit rate is to be expected; the reader should also bear in mind that
even when parsing all \numberOFlogFILES{} log files not all the rules are
exercised (some of the rules are for parsing log lines which only appear in
other log files).

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly (many of the
complications discussed in \sectionref{additional complications} were
detected in this way).  The parser issues warnings when it detects any
errors, some of which may alert the user to a problem, e.g.\ when a queueid
is reused before the previous mail is fully dealt with, when a queueid or
\pid{} is not found in the state tables,\footnote{There will often be
warnings about a missing queueid or \pid{} in the first few hundred or
thousand log lines because the earlier log lines for those connections or
mails are in the previous log file; loading the saved state from the
previous log file will solve this problem.} or when there are problems
tracking a child mail (see \sectionref{tracking re-injected mail}).  There
should be few or no warnings when parsing, and when finished parsing the
state table should only contain entries for mails which had yet to be
delivered when the log files ended, or were accepted before the log files
began.

@@

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails which
are awaiting delivery after the period covered by the log files, though
they may also contain mails whose initial entries are not contained in the
log files.  Any other entries are evidence of a failure in parsing or an
aberration in the log files.  After parsing the \numberOFlogFILES{} test
log files the state tables contain 18 entries, breaking down into:

@@

    \item 1 connection which started only seconds before the log files
        ended and had not yet been fully transferred from client to server.

    \item 1 mail which had been accepted only seconds before the log files
        ended and had not yet been delivered.

@@

    \item 7 mails which had yet to be delivered due to repeated failures.

@@

    \item When dealing with the complications described in
        \sectionref{complications} and \sectionref{additional
        complications} the solutions are as specific and restrictive as
        possible, with the goal of minimising the number of false
        positives.  In addition the solution to the \textit{Out of order
        log files\/} complication described in \sectionref{out of order log
        lines} imposes conditions which each reassembled mail must comply
        with to be acceptable.

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log and a dump of the resulting database,
verifying that for each of the input lines the parser used the correct rule
and executed the correct action, which in turn produced the correct result
and inserted the correct data in the database.  The log segment produced 4
warnings, 10 mails remaining in the state tables, and 1625 connections
correctly entered in the database.

>>>>

XXX COMPARE PARSING + ACTIONS TO JUST PARSING\@.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day; \graphref{Mails received per day} shows that as expected there are far
more mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \SMTP{} only; in particular the mail loops noticeable in
other graphs do not contribute to these figures.

@@

\Graphref{parsing time vs file size vs number of lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline
\tab{}\texttt{$\backslash$username, username@domain} \newline This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Ordinarily this works well, but unfortunately
in this instance the user noticed they had not received any mail in a while
and opted to send a test mail to themselves, causing a loop not caught by
Postfix:

@@

    \item Postfix accepts the mail for username@domain, but during delivery
        it notices that the \texttt{Delivered-To} header already contains
        the address it is currently delivering to, and therefore sends a
        bounce notification with sender address \texttt{<>} to the original
        sender: username@domain.

@@

    \item Postfix accepts the forwarded bounce notification but during
        deliver the mail it notices that the \texttt{Delivered-To}
        header already contains the address currently being delivered to,
        and sends a bounce notification to the sender: username@domain.

@@

The different distribution of log lines during a mail forwarding loop has
unexpected effects on parsing time, described in detail in
\sectionref{rule ordering for efficiency}, \sectionref{scalability as the
number of rules rises}, and \sectionref{Caching compiled regexes}.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% match the vast majority of log lines, with the remaining log lines
split across the other 90\% of the rules (as shown in \graphref{rule hits
graph}).  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently matching rules before those that match less frequently.  To
test this hypothesis three full test runs were performed with different
rule orderings:

@@

    \item [optimal]  The most optimal order, according to the hypothesis:
        rules that match most often will be tried first.

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general the
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of lines versus parsing time in
\graphref{parsing time vs file size vs number lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

@@

    \item Check if there are any rules for the Postfix component that
        produced the log line; if not then skip the log line.

@@

    \item Each Postfix component whose log lines are of interest must have
        at least one rule or its log lines will be silently skipped; in the
        extreme case of zero rules the parser would happily skip every log
        line.  There may be any number of log lines from other programs
        intermingled in the log file, and there are some Postfix programs
        that do not produce any log lines of interest.

    \item There must be a rule to match each different log line produced by
        each program; if a log line is not successfully matched the parser
        will issue a warning.  Rules should be as specific and tightly
        bound as possible to ensure accurate parsing:\footnote{A rule that
        matches zero or more of any character will successfully parse every
        log line, but not in a meaningful way.} most log lines contain
        fixed strings and have a rigid pattern, so this is not a problem.

@@

Warnings are issued for any log lines that are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6,039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.\footnote{Each log
line was examined and the correct \regex{} identified from the
\numberOFrules{} rules in the database; the correct \regex{} was then
compared to the \regex{} that was used by the parser.}  The sample was
generated by running the following command:

@@

\noindent{}to randomly extract roughly one line in every 10,000.  Although
on initial appearances exercising only 36 rules (from a total of
\numberOFrules{}) when parsing 6039 log lines seems quite low, after
examining \graphref{rule hits graph} it becomes apparent that such
a low hit rate is to be expected; the reader should also bear in mind that
even when parsing all \numberOFlogFILES{} log files not all the rules are
exercised (some of the rules are for parsing log lines that only appear in
other log files).

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails that were not handled properly (many of the
complications discussed in \sectionref{additional complications} were
detected in this way).  The parser issues warnings when it detects any
errors, some of which may alert the user to a problem, e.g.\ when a queueid
is reused before the previous mail is fully dealt with, when a queueid or
\pid{} is not found in the state tables,\footnote{There will often be
warnings about a missing queueid or \pid{} in the first few hundred or
thousand log lines because the earlier log lines for those connections or
mails are in the previous log file; loading the saved state from the
previous log file will solve this problem.} or when there are problems
tracking a child mail (see \sectionref{tracking re-injected mail}).  There
should be few or no warnings when parsing, and when finished parsing the
state table should only contain entries for mails that had yet to be
delivered when the log files ended, or were accepted before the log files
began.

@@

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails that
are awaiting delivery after the period covered by the log files, though
they may also contain mails whose initial entries are not contained in the
log files.  Any other entries are evidence of a failure in parsing or an
aberration in the log files.  After parsing the \numberOFlogFILES{} test
log files the state tables contain 18 entries, breaking down into:

@@

    \item 1 connection that started only seconds before the log files
        ended and had not yet been fully transferred from client to server.

    \item 1 mail that had been accepted only seconds before the log files
        ended and had not yet been delivered.

@@

    \item 7 mails that had yet to be delivered due to repeated failures.

@@

    \item When dealing with the complications described in
        \sectionref{complications} and \sectionref{additional
        complications} the solutions are as specific and restrictive as
        possible, with the goal of minimising the number of false
        positives.  In addition the solution to the \textit{Out of order
        log files\/} complication described in \sectionref{out of order log
        lines} imposes conditions that each reassembled mail must comply
        with to be acceptable.

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log and a dump of the resulting database,
verifying that for each of the input lines the parser used the correct rule
and executed the correct action, that in turn produced the correct result
and inserted the correct data in the database.  The log segment produced 4
warnings, 10 mails remaining in the state tables, and 1625 connections
correctly entered in the database.

### 2008-08-25 doc/logparser.tex
Part two of the great renaming complete. Graphs still need to be renamed.
Renamed actions because Carl can't understand that the IGNORE action doesn't mean that the log line wasn't parsed, so other people might have the same problem:     IGNORE              => UNINTERESTING     REJECTION           => DELIVERY_REJECTED     SAVE_BY_QUEUEID     => SAVE_DATA
Removed some unused labels cluttering the paper. Minor cleanup.
Final suggestions from Carl. Remaining corrections from my proofreading. Resolved all XXX items. Improved bibliography entries.
Changes made following suggestions from diction.

@@

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding the log files, and thus accuracy in reconstructing the
    actions taken by Postfix to generate the log files.  This paper
    describes the creation of a parser, documenting the parsing algorithm,
    explaining the difficulties encountered and the solutions developed,
    with reference to an implementation which stores data gleaned from the
    log files in an SQL database.  The gathered data can then be used to
    optimise current anti-spam measures, provide a baseline to test new
    anti-spam measures against, or to produce statistics showing how
    effective those measures are.

@@

XXX NEED A NAME FOR THE ARCHITECTURE, AND A DIFFERENT NAME FOR THE
IMPLEMENTATION\@.

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-other-parsers-review.tex}
\input{logparser-architecture.tex}
\input{logparser-implementation.tex}
\input{logparser-parsing-rules.tex}
\input{logparser-results.tex}
\input{logparser-conclusion.tex}

>>>>

%\setlength{\floatsep}{0pt}
%\setlength{\intextsep}{0pt}
%\setlength{\textfloatsep}{0pt}

@@

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding the log files, and thus accuracy in reconstructing the
    actions taken by Postfix to generate the log files.  This paper
    describes the creation of a parser, documenting the parsing algorithm,
    explaining the difficulties encountered and the solutions developed,
    with reference to an implementation that stores data gleaned from the
    log files in an SQL database.  The gathered data can then be used to
    optimise current anti-spam measures, provide a baseline to test new
    anti-spam measures against, or to produce statistics showing how
    effective those measures are.

@@

XXX NEED A NAME FOR THE ARCHITECTURE, AND A DIFFERENT NAME FOR THE
IMPLEMENTATION\@.  Parse::FAR\@???

XXX CHECK ALL ``UNINTERESTING'' RULES TO SEE IF THEY COULD BE USED\@.

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-other-parsers-review.tex}
\input{logparser-architecture.tex}
\input{logparser-implementation.tex}
\input{logparser-results.tex}
\input{logparser-conclusion.tex}
